---
title: "Clustering Player and Ball Coordinates"
date: "30/05/2020"
output: md_document
---

This is a very very very WIP post.

## Data Handling

- I take one in every three frames just to lower the number of frmaes that need to be managed. This means there are aabout 8 frames per second which is still sufficient resolution for us to run this analysis.

- Players substituted for each other are treated as the same player for purpose of clustering, analysis, etc. This is not a great assumption because players frequently substitute players who weren't playing in the same role but I don't have a better simple way to handle this yet. You could have an addition player role role assignment sort of model, of which there are a few, to assign the correct roles and then cluster.

## Procedure

- This post focuses on the home team attacking during sample game 2.

- Euclidean distance calculated between the players or the ball in each frame and passed to a hierarchical clustering algorithm.

- I extract two levels of clusters from this, a set of 6 clusters and set of 48 clusters. I call the first set of clusters, cluster groups, and the second set of clusters just clusters. No strong reason behind why six clusters except it's a good balance between splitting the data into an unmanageably large number of clusters and not splitting it enough and being unable to infer anything out of them. No strong reason behind 48 either except it's an easy number to work with.

- Also tried clustering just the ball locations and clustering player locations leaving out the GK and translating the rest of the players such that the mean location of all the player in that frame was at (0,0). That didn't look as good as this.

```{r Setup, cache=FALSE, echo=FALSE, warning=FALSE, message=FALSE, fig.width = 15, fig.height = 8, results = 'hide'}

rm(list = ls())

library(CodaBonito)
library(data.table)
library(ggplot2)
library(ggrepel)
library(scales)
library(zoo)
library(clue)
library(knitr)
theme_set(theme_bw(12))
opts_chunk$set(
   dev.args=list(
      bg="transparent"
   ),
   dpi = 300
)

theme_pitch_override = function() {
      
   # http://www.sthda.com/english/wiki/ggplot2-themes-and-background-colors-the-3-elements
   
   theme(
      plot.background = element_rect(fill = 'black'), 
      panel.background = element_rect(fill = 'black', color = '#666666'),
      legend.title = element_text(color = 'white'),
      legend.text = element_text(color = 'white', size = rel(0.8)),
      plot.title = element_text(color = 'white', size = rel(1.2)),
      strip.background = element_rect(fill = "#666666", colour = NA),
      strip.text = element_text(colour = "#cccccc", size = rel(0.8)),

   )
   
}
 
theme_non_pitch_override = function() {
   theme_pitch_override() +
   theme(
      panel.grid.major = element_line(colour = "#333333"), 
      panel.grid.minor = element_line(colour = "#333333", size = 0.25), 
      axis.line = element_blank(), 
      axis.text = element_text(
         size = rel(0.8), 
         colour = "#cccccc"
      ),
      axis.text.x = element_text(
         vjust = 1
      ),
      axis.text.y = element_text(
         hjust = 1
      ),
      axis.ticks = element_line(colour = "#111111"), 
      axis.title.x = element_text(
         colour = "#cccccc"
      ), 
      axis.title.y = element_text(
         angle = 90, 
         colour = "#cccccc"
      )
   )
   
}

# Instructions
# You would have downloaded and saved the data from Github at some location on
# your system. In my case, it's /media/ask/Data/Personal/Projects/Personal/sample-data
# Replace the location below with the respective location from your machine
cDataRootFolder = '/media/ask/Data/Personal/Projects/Personal/sample-data/data/'
cGameName = 'Sample_Game_2'
cTeam = 'Home'
nXLimit = 120
nYLimit = 80
# upper limit of speed to cap weird speed values
nUpperLimitSpeed = 5 * 120 / 105
cOutputFolderRoot = '/media/ask/Data/Personal/Projects/Personal/MakingFriendsWithTrackingData/'


dtPlayerLabels = data.table(
   Player = c(
     paste0('HomePlayer', 1:11)
   ),
   Label = c(
     'RB','RCB','LCB','LB','RW','RCM','LCM','LW','SS','CF','GK'
   )
)

dtPlayerLabels[,
   Label := factor(
      Label,
      levels = c('GK', 'RB','RCB','LCB','LB','RW','RCM','LCM','LW','SS','CF'),
      ordered = T
   )
]

```

```{r DataLoading, cache=FALSE, echo=FALSE, warning=FALSE, message=FALSE, fig.width = 15, fig.height = 8, results = 'hide'}

# loading the data in
lData = fParseTrackingDataBothTeams(
   cRootPath = cDataRootFolder,
   cGameName = cGameName,
   nXLimit = nXLimit,
   nYLimit = nYLimit,
   xMaxBB = 1,
   yMaxBB = 1,
   nUpperLimitSpeed = nUpperLimitSpeed
)

dtPlayerInOuts = lData$dtTrackingData[
   Tag != 'Ball', 
   list(
      StartFrame = min(Frame), 
      EndFrame = max(Frame)
   ), 
   list(Tag, Player)
][
   !(StartFrame == 1 & EndFrame == max(EndFrame))
]

dtSubstitutions = merge(
    dtPlayerInOuts[, list(Tag, PlayerOut = Player, Frame = EndFrame)],
    dtPlayerInOuts[, list(Tag, PlayerIn = Player, Frame = StartFrame)],
    c('Tag', 'Frame')
)

setorder( dtSubstitutions, Frame )

lData$dtTrackingData[, Player2 := as.character(NA)]

for ( i in seq(nrow(dtSubstitutions)) ) {
   
   cNewPlayerName = lData$dtTrackingData[
      Player %in% dtSubstitutions[i, c(PlayerOut, PlayerIn)],
   ][
      !is.na(Player2),
      Player2[1]
   ]
   
   if ( is.na(cNewPlayerName) ) {
      
      cNewPlayerName = paste0(
         dtSubstitutions[i, Tag], 
         'Player', 
         dtSubstitutions[i, paste(gsub(x = c(PlayerOut,PlayerIn), pattern = '.*Player', replacement = ''), collapse = '.')]
      )
      
   }
   
   lData$dtTrackingData[
      Player %in% dtSubstitutions[i, c(PlayerOut, PlayerIn)],
      Player2 := cNewPlayerName
   ]
   
   dtPlayerLabels[
      Player %in% dtSubstitutions[i, c(PlayerOut, PlayerIn)],
      Player2 := cNewPlayerName
   ]
   
}

lData$dtTrackingData[
   !is.na(Player2),
   Player := Player2
]

dtPlayerLabels[
   !is.na(Player2),
   Player := Player2
]

lData$dtTrackingData[,
   Player2 := NULL
]

dtPlayerLabels[,
   Player2 := NULL
]

```
 
 
```{r DataPrepAttack, cache=FALSE, echo=FALSE, warning=FALSE, message=FALSE, fig.width = 15, fig.height = 8, results = 'hide'}

viTrackingFrames = lData$dtTrackingData[Player == 'Ball', unique(Frame)]
viTrackingFrames = viTrackingFrames[viTrackingFrames %% 3 == 0]
viTrackingFrames = viTrackingFrames[
   viTrackingFrames < dtPlayerInOuts[
      Tag == cTeam, 
      min(StartFrame[StartFrame != 1])
   ]
]
# inferring which team was in control of the ball
dtAttackingTeam = merge(
   lData$dtEventsData[
       Type %in% c("PASS", "SHOT", "SET PIECE", "RECOVERY") |
       Subtype %in% c("GOAL KICK", "KICK OFF"),
       list(AttackingTeam = Team[1]),
       list(Frame = StartFrame, EndFrame)
   ],
   data.table(Frame = viTrackingFrames),
   'Frame',
   all = T
)

dtAttackingTeam[is.na(EndFrame), EndFrame := Frame]

dtAttackingTeam = dtAttackingTeam[,
   .SD[which.max(EndFrame)],
   list(Frame)
][,
   AttackingTeam := na.locf(AttackingTeam, na.rm = F)
]

dtAttackingTeam = dtAttackingTeam[
   Frame %in% viTrackingFrames,
   list(Frame, AttackingTeam)
]

lData$dtTrackingData = lData$dtTrackingData[
   Frame %in% viTrackingFrames
]

```


 
```{r DistanceCalculation, cache=FALSE, echo=FALSE, warning=FALSE, message=FALSE, fig.width = 15, fig.height = 8, results = 'hide'}
WhatTypeOfDistance = 'Absolute'

cOutputFolder = paste0(
   cOutputFolderRoot,
   '/DistanceMatrices/', WhatTypeOfDistance
)


dir.create(
   cOutputFolder,
   showWarnings = F,
   recursive = T
)

if ( T ) {
      
   iFrameBucketSpan = 300
   
   dtDistances = lapply(
      rev(seq(ceiling(length(viTrackingFrames)/iFrameBucketSpan))),
      function ( iFrameWindowIndex ) {
         
         for ( cAttackingTeam in dtAttackingTeam[, unique(AttackingTeam)] ) {
            
            for ( cTag in lData$dtTrackingData[, unique(Tag)] ) {
                  
               dir.create(
                  paste0(cOutputFolder, '/', cTag, '/', cAttackingTeam),
                  showWarnings = F,
                  recursive = T
               )
               
               cFileName = paste0(
                  cOutputFolder, '/', cTag, '/', cAttackingTeam, '/', iFrameWindowIndex,'.Rdata'
               )
               
               if ( !file.exists(cFileName) ) {
                  
                  print(Sys.time())
                  print(cTag)
                  print( iFrameWindowIndex / ( length(viTrackingFrames) / iFrameBucketSpan) )
                  
                  viFrames = viTrackingFrames[(((iFrameWindowIndex - 1) * iFrameBucketSpan ) + 1):( (iFrameWindowIndex) * iFrameBucketSpan )]
                  viFrames = dtAttackingTeam[
                     AttackingTeam == cAttackingTeam,
                     intersect(
                        Frame, viFrames
                     )
                  ]
                  
                  if ( length(viFrames) > 0 ) {
                     
                     dtTrackingDataSubset = lData$dtTrackingData[ 
                        Frame %in% viFrames 
                     ][
                        Tag %in% c(cTag)
                     ][, 
                        list(Tag, Frame, Player, X, Y)
                     ]
                     
                     dtDistances = merge(
                        dtTrackingDataSubset,
                        lData$dtTrackingData[
                           Tag %in% c(cTag)
                        ][
                           Frame <= max(viFrames)
                        ][
                           Frame %in% dtAttackingTeam[
                              AttackingTeam == cAttackingTeam,
                              Frame
                           ]
                        ][, 
                           list(Frame, Player, X, Y)
                        ],
                        c('Player'),
                        allow.cartesian = T
                     )[
                        Frame.x >= Frame.y
                     ]
                     
                     if ( nrow(dtDistances) > 0 ) {
                        
                        print(dtDistances[, range(Frame.x)])
                        
                        dtDistances[,
                           Distance := 
                              ( ( X.x - X.y ) ^ 2 ) +
                              ( ( Y.x - Y.y ) ^ 2 )
                        ]
                        
                        dtDistances = dtDistances[, 
                           list(
                              Distance = sum(Distance) ^ 0.5,
                              Items = .N
                           ),
                           list(Frame.x, Frame.y)
                        ]
                        
                        gc()
                           
                     } else {
                        
                        dtDistances = data.table()
                     
                     }
                     
                  } else {
                     
                     dtDistances = data.table()
                  
                  }
                  
                  save(
                     list = 'dtDistances',
                     file = cFileName
                  )
                  
               }
               
            }
         }
         
         NULL
         
      }
   )
   
   rm(dtDistances)
   
}


```


 
```{r DataPostProcessAttack, cache=FALSE, echo=FALSE, warning=FALSE, message=FALSE, fig.width = 15, fig.height = 8, results = 'hide'}

if ( T ) {
   
   cTagsToCapture = c(cTeam, 'Ball')
   cAttackingTeam = cTeam
   
   vcFiles = c(
      list.files(paste0(cOutputFolder, '/',cTeam,'/', cAttackingTeam), recursive = T)
   )
   
   cFileName = paste0('/media/ask/Data/Personal/tmp/trackingclustering.', WhatTypeOfDistance, '.', cGameName, '.', cAttackingTeam, '.', paste(sort(cTagsToCapture), collapse = ''), '.Rdata')
   
   if ( file.exists(cFileName) ) {
      
      load(cFileName)
      
   } else {
      
      dtDistances = rbindlist(
         lapply(
            vcFiles,
            function(cFile) {
               
               if ( interactive() ) {
                  print(cFile)
               }
               
               dtDistances = rbindlist(
                  lapply(
                     cTagsToCapture,
                     function (cTagToCapture) {
                        
                        load(paste0(cOutputFolder, '/', cTagToCapture, '/', cAttackingTeam, '/', cFile))
                        
                        if ( nrow(dtDistances) > 0 ) {
                           
                           # print(head(dtDistances))
                           
                           if ( cTagToCapture %in% c('Home','Away') ) {
                              dtDistances[, Distance := ( 11 * ( Distance ^ 2 ) / Items ) ^ 0.5]
                              dtDistances[, Items := 11]
                           }
                           
                           dtDistances[, Tag := cTagToCapture]
                           
                           dtDistances = dtDistances[
                              Frame.x %in% viTrackingFrames
                           ][
                              Frame.y %in% viTrackingFrames
                           ]
                           
                        } else {
                           
                           dtDistances = data.table()
                           
                        }
                        
                        dtDistances
                        
                     }
                  ),
                  fill = T
               )
               
               if ( nrow( dtDistances ) > 0 ) {
                  
                  dtDistances = dtDistances[, sum(Distance^2)^0.5, list(Frame.x, Frame.y)]
                  
                  dtDistances = dcast(
                     dtDistances,
                     Frame.x ~ Frame.y,
                     fun.aggregate = mean,
                     value.var = 'V1'
                  )
                  
               }
               
               dtDistances
               
            }
            
         ),
         fill = T
      )
      
      save(
         list = 'dtDistances',
         file = cFileName
      )
      
   }
   
   
   dtDistances = dtDistances[Frame.x %in% viTrackingFrames]
   dtDistances = dtDistances[, c(intersect(colnames(dtDistances), viTrackingFrames), 'Frame.x'), with = F]
   dtDistances = dtDistances[order(as.integer(Frame.x))]
   
   setcolorder(
      dtDistances,
      c('Frame.x', dtDistances[, Frame.x])
   )
   
   if ( F ) { 
      
      cAttackingTeam = 'Home'
      for ( cTagToCapture in c('Home','Away','Ball') ) {
         
          for ( cFile in vcFiles ) {
              if(file.exists(paste0(cOutputFolder, '/', cTagToCapture, '/', cAttackingTeam, '/', cFile))) {
          load(paste0(cOutputFolder, '/', cTagToCapture, '/', cAttackingTeam, '/', cFile))
          if ( 'MeanDistance' %in% colnames(dtDistances) ) {
              print(cFile)
              file.remove(paste0(cOutputFolder, '/', cTagToCapture, '/', cAttackingTeam, '/', cFile))
          }
          rm(dtDistances)
              }
          }
      }
            
   }
   
   if ( F ) {
      
      for ( i in 1:nrow(dtDistances) ) {
          if ( any(is.na(
             dtDistances[i,2:(i+1)]
          )) ) stop()
      }
      
   }
   
}

```


 
```{r ClusteringAttack, cache=FALSE, echo=FALSE, warning=FALSE, message=FALSE, fig.width = 15, fig.height = 8, results = 'hide'}

if ( F ) {
   
   set.seed(1)
      
   kmeansInteractions = kmeans(
      dtSocialDistancingWide[, 
         !c(
            'Frame',
            # 'Ball.x','Ball.y',
            ''
         ), 
         with = F
      ], 
      centers = 48
   )
   
   dtSocialDistancingWide[, Cluster := cl_predict(kmeansInteractions)]

}

if ( T ) {
   
   hcModel = hclust(
      # dist(dtSocialDistancingWide[
      #    Frame == min(Frame) | Frame %% 2 == 0, 
      #    !c('Frame','Stretch','Cluster')
      # ]),
      as.dist(
         dtDistances[
            # Frame.x <= 95703
         ][, !'Frame.x', with = F][
            # 1:.N, 
            # 1:dtDistances[, max(which(Frame.x <= 95703))]
         ],
         diag = T
      ),
      method = 'ward.D2'
   )
   
   dtClusterResults = data.table(
      Frame = dtDistances[
         # Frame.x <= 95703
         , Frame.x
      ],
      # ClusterGroup := cutree(hcModel, k = 1 + max(which((diff(rev(hcModel$height)[1:50])) < -500)))
      ClusterGroup = cutree(hcModel, k = 6), 
      Cluster = cutree(hcModel, k = 48)
   )[,
      ClusterGroup.Cluster := paste0(ClusterGroup, '.', formatC(Cluster, width = 2, flag = '0'))
   ]
   
   setorder(
      dtClusterResults,
      Frame
   )
   
   dtClusterResults[, Instance := cumsum(c(1, diff(Frame) > 100)), Cluster]
   
   rm(dtDistances)
   
   # dtSocialDistancingWide[, Cluster := na.locf(Cluster, na.rm = F)]

}



```

## Understanding cluster groups
 
```{r PrintingClusterGroups, cache=FALSE, echo=FALSE, warning=FALSE, message=FALSE, fig.width = 10, fig.height = 6}

dtClusterGroupOccurrence = dtClusterResults[, 
   list(
      Instance = length(unique(Instance)), 
      .N
   ), 
   list(
      ClusterGroup
   )
][, 
   list(ClusterGroup, Instance, Time_pct = round(N / sum(N), 2) * 100)
]

p1 = ggplot(
   merge(
      dtClusterResults[!is.na(Cluster)],
      lData$dtTrackingData[Tag == 'Ball', list(Frame, X, Y)],
      'Frame'
   )
) +
   geom_pitch(cLineColour = '#333333',cPitchColour = 'black') + 
   geom_point(
      aes(x = X, y = Y),
      color = 'white',
      size = 0.1
   ) +
   labs(
      title = 'Where the ball was in each cluster'
   ) +
   scale_color_discrete(guide = F)
   
vcFacetLabels = dtClusterGroupOccurrence[, paste0('Cluster group ', ClusterGroup,'\n', Time_pct,'% of the total time, ', Instance, ' instance(s)')]
names(vcFacetLabels) = as.character(dtClusterGroupOccurrence[, ClusterGroup])

p1 = p1 + 
   # geom_text(
   #    data = dtClusterGroupOccurrence,
   #    aes(
   #       x = 0,
   #       y = -5,
   #       label = paste0(Time_pct,'% of the total time, ', Instance, ' instance(s)')
   #    ),
   #    size = 4,
   #    hjust = 0,
   #    vjust = 1,
   #    color = 'white'
   # ) +
   facet_wrap(
      ~ClusterGroup, 
      labeller = labeller(
         ClusterGroup = vcFacetLabels
      )
   ) +
   theme_pitch() + 
   theme_pitch_override()

print(p1)

```

 
The six clusters broadly capture situations which are in place when the ball is in various parts of the field. 
Clusters 1 and 3 capture the two sides in the defensive half. Clusters 2 and 5 capture the two sides in the 
offensive half. Cluster 4 is when the ball is in midfield. Cluster 6 looks like the ball is in similar areas as
cluster 4 so the difference between the two might be interesting.

## Understanding clusters

Diving into the smaller clusters -
 
```{r PrintingRresultsAttack, cache=FALSE, echo=FALSE, warning=FALSE, message=FALSE, fig.width = 10, fig.height = 12}
# plot 12 clusters at a time in 4 batches

HowManyAtATime = 12

for ( iPart in 1:4 ) {
   
   dtChartData = merge(
     lData$dtTrackingData[Tag %in% c('Ball', cTeam)],
     dtClusterResults[, list(Frame, ClusterGroup, Cluster, ClusterGroup.Cluster, Instance)],
     'Frame'
   )
   
   dtChartData = merge(
      dtChartData,
      dtPlayerLabels,
      'Player',
      all.x = T
   )
   
   dtChartData[Player != 'Ball', Player := Label]
   dtChartData[, Label := NULL]
   
   dtChartDataOccurrence = dtChartData[, 
      list(
         Instance = length(unique(Instance)), 
         .N
      ), 
      list(
         ClusterGroup, Cluster, ClusterGroup.Cluster
      )
   ]
   
   dtChartDataOccurrence[, 
      Time_pct := round(N / sum(N), 2) * 100
   ]
   
   if ( F ) {
      
      dtChartDataOccurrence[, Cluster2 := 1 + .N - rank(Instance, ties.method = 'random')]
      
      dtChartData = merge(
         dtChartData,
         dtChartDataOccurrence[, list(Cluster, Cluster2)],
         'Cluster'
      )
      
      dtChartData[, Cluster := Cluster2]
      dtChartData[, Cluster2 := NULL]
      dtChartDataOccurrence[, Cluster := Cluster2]
      dtChartDataOccurrence[, Cluster2 := NULL]
      
   }
   
   dtChartData = dtChartData[,
    .SD[Frame %in% sample(unique(Frame), min(length(unique(Frame)), 10))],
    Cluster
   ]
   
   if ( F ) {
         
      dtChartData[,
         X := X - X[Player == 'Ball'],
         Frame
      ]
      
      dtChartData[,
         Y := Y - Y[Player == 'Ball'],
         Frame
      ]
      
   }
   
   dtChartDataHull = dtChartData[,
     .SD[chull(X,Y)],
     list(Player, Cluster)
   ]
   
   
   
   
   
   setorder(
      dtChartDataOccurrence,
      ClusterGroup,
      Cluster
   )
   
   dtChartDataOccurrence = dtChartDataOccurrence[(c(1:HowManyAtATime) + ((iPart-1) * HowManyAtATime))]
   dtChartData = dtChartData[Cluster %in% dtChartDataOccurrence[, Cluster]]
   
   vcFacetLabels = dtChartDataOccurrence[, paste0('Cluster ', ClusterGroup.Cluster,'\n', Time_pct,'% of the total time, ', Instance, ' instance(s)')]
   names(vcFacetLabels) = as.character(dtChartDataOccurrence[, ClusterGroup.Cluster])

   p1 = ggplot(
     dtChartData
   ) + 
      geom_pitch(cLineColour = '#333333', cPitchColour = 'black') + 
     # geom_hline(yintercept = 0) +
     # geom_vline(xintercept = 0) +
     # geom_polygon(
     #    data = dtChartDataHull[Tag != 'Ball'],
     #    aes(
     #      x = X,
     #      y = Y,
     #      group = Player,
     #      color = Player
     #    ),
     #    fill = NA,
     #    alpha = 0.1
     #  ) + 
      # geom_point(
      #   aes(
      #     x = X,
      #     y = Y,
      #     color = Player
      #   ),
      #   alpha = 0.1
      # ) +
      # geom_density_2d(
      #   data = dtChartData[
      #      Tag != 'Ball'
      #   ],
      #   aes(
      #     x = X,
      #     y = Y,
      #     color = Player,
      #     group = paste(Player, Cluster)
      #   ),
      #   size = 2
      # ) +
      # geom_point(
      #   data = dtChartData[
      #      Tag != 'Ball'
      #   ][, list(X = mean(X), Y = mean(Y)), list(Cluster, Player)],
      #   aes(
      #     x = X,
      #     y = Y
      #     
      #   ),
      #   size = 4
      # ) +
      # geom_point(
      #   data = dtChartData[
      #      Tag != 'Ball'
      #   ][, list(X = mean(X), Y = mean(Y)), list(Cluster, Player)],
      #   aes(
      #     x = X,
      #     y = Y,
      #     # color = Player
      #   ),
      #   size = 3
      # ) +
      geom_point(
        data = dtChartData[
           Tag == 'Ball'
        ][, list(X = mean(X), Y = mean(Y)), list(ClusterGroup, Cluster, ClusterGroup.Cluster, Player)],
        aes(
          x = X,
          y = Y
        ),
        color = 'red',
        size = 2
      ) +
      geom_text(
        data = dtChartData[
           Tag != 'Ball'
        ][, list(X = mean(X), Y = mean(Y)), list(ClusterGroup, Cluster, ClusterGroup.Cluster, Player = gsub(Player, pattern = '.*Player', replacement = ''))],
        aes(
          x = X,
          y = Y,
          label = Player
        ),
        color = 'red',
        size = 3
      ) +
      # geom_text(
      #    data = dtChartDataOccurrence,
      #    aes(
      #       x = 0,
      #       y = -5,
      #       label = paste0(Time_pct,'% of the total time, ', Instance, ' instance(s)')
      #    ),
      #    color = 'red',
      #    size = 4,
      #    hjust = 0
      # ) +
      # facet_grid(Player~Cluster) +
      scale_colour_discrete(guide = F) +
      facet_wrap(
         ~ClusterGroup.Cluster, 
         nrow = 4, 
         labeller = labeller(
            ClusterGroup.Cluster = vcFacetLabels
         )
      ) +
      geom_blank() + 
      coord_fixed() +
      labs(
         # title = paste0('Positions of the ', cTeam, ' players and ball in game ', cGameName,' while the ', dtAttackingTeam[Frame %in% viTrackingFrames, AttackingTeam[1]], ' team is attacking - ', iPart),
         title = paste0('Average positions of the players and ball in each cluster - ', iPart),
         subtitle = 'Each box shows the average position of the players or the ball in the cluster.'
      ) +
      theme_pitch() + 
      theme_pitch_override()
   
   print(p1)

}

```

## Some things that you can quickly see

Note: This section hasn't been updated for the new clusters so the example clusters are probably wrong but the observations should still hold

- They seem to either play in a 4-4-1-1 / 4-4-2, see clusters 4, 5, 7, 8

- Or they switch to a back three sometimes with the two CBs and either the LB or the RCM, see clusters 1, 2, 3 

- The LCM is more involved further up the field than the RCM, see clusters 2, 3, 9

- The RB has more of an attacking role than the LB, see clusters 1, 6, 9

- Attacks go through the right more often than the left

- The LB has a more defensive role

- The RW has a more central role, either as an RWF or a  CM, see clusters 1, 2, 3

- The SS, RW, and RB have a lot of combinations, see clusters 6, 9

- The LW, and the RB as well, try to provide width, see clusters 1, 3, 6a

- Clusters 16, 30 might be corners
 
## Exploring a cluster in more detail

```{r DetailedLookAtOneCluster, cache=FALSE, echo=FALSE, warning=FALSE, message=FALSE, fig.width = 15, fig.height = 10, results = 'hide'}

iSelectedCluster = dtClusterResults[!is.na(Cluster), .N, Cluster][which.max(N), Cluster]

dtChartData = merge(
  lData$dtTrackingData[Tag %in% c('Ball', cTeam)],
  dtClusterResults[!is.na(Cluster)][iSelectedCluster == Cluster, list(Frame, Cluster, Instance)],
  'Frame'
)

dtChartData = merge(
   dtChartData,
   dtPlayerLabels,
   'Player',
   all.x = T
)

dtChartData[Player != 'Ball', Player := Label]
dtChartData[, Label := NULL]
   
if ( F ) {
   
   dtChartDataOccurrence = dtChartData[, 
      list(
         Instance = length(unique(Instance)), 
         .N
      ), 
      list(
         ClusterGroup, Cluster, ClusterGroup.Cluster
      )
   ][, 
      list(ClusterGroup, Cluster, ClusterGroup.Cluster, Instance, Time_pct = round(N / sum(N), 2) * 100)
   ]
   
   dtChartDataOccurrence[, Cluster2 := 1 + .N - rank(Instance, ties.method = 'random')]
   
   dtChartData = merge(
      dtChartData,
      dtChartDataOccurrence[, list(Cluster, Cluster2)],
      'Cluster'
   )
   
   dtChartData[, Cluster := Cluster2]
   dtChartData[, Cluster2 := NULL]
   dtChartDataOccurrence[, Cluster := Cluster2]
   dtChartDataOccurrence[, Cluster2 := NULL]
   
}


if ( F ) {
      
   dtChartData[,
      X := X - X[Player == 'Ball'],
      Frame
   ]
   
   dtChartData[,
      Y := Y - Y[Player == 'Ball'],
      Frame
   ]
   
}

dtChartDataHull = dtChartData[,
  .SD[chull(X,Y)],
  list(Player, Cluster)
]

   
   
   
p1 = ggplot(
  dtChartData
) +
   geom_pitch(cLineColour = '#333333',cPitchColour = 'black') + 
   # geom_point(
   #   aes(
   #     x = X,
   #     y = Y
   #   ),
   #   alpha = 0.1
   # ) +
   geom_path(
     aes(
       x = X,
       y = Y,
       group = paste(Instance, Cluster, Player)
     ),
     color = 'white'
   ) +
   geom_point(
     data = dtChartData[
     ][, list(X = mean(X), Y = mean(Y)), list(Cluster, Player = gsub(Player, pattern = '.*Player', replacement = ''))],
     aes(
       x = X,
       y = Y
     ),
     size = 3,
     color = 'red'
   ) +
   geom_point(
     data = dtChartData[
     ][, .SD[which.max(Frame)], list(Player, Cluster, Instance)],
     aes(
       x = X,
       y = Y
     ),
     color = 'white'
   ) +
   facet_wrap(~Player) +
   scale_colour_discrete(guide = F) +
   geom_blank() + 
   theme_pitch() + 
   theme_pitch_override() +
   coord_fixed() +
   labs(
      title = paste0('Detailed positions of the ', cTeam, ' players and ball in cluster ', dtClusterResults[Cluster == iSelectedCluster, ClusterGroup.Cluster[1]]),
      subtitle = 'Mean position in red, all positions during the game in that cluster in black.'
   )

print(p1)


```


## Exploring a player's role in more detail

```{r DetailedLookAtOnePlayer, cache=FALSE, echo=FALSE, warning=FALSE, message=FALSE, fig.width = 15, fig.height = 17, results = 'hide'}

cSelectedPlayer = dtPlayerLabels[Label == 'RB', Player]

dtChartData = merge(
  lData$dtTrackingData[Player %in% c(cSelectedPlayer, 'Ball')][Tag %in% c('Ball', cTeam)],
  dtClusterResults[!is.na(Cluster)][, list(Frame, ClusterGroup, Cluster, ClusterGroup.Cluster, Instance)],
  'Frame'
)

dtChartData = merge(
   dtChartData,
   dtPlayerLabels,
   'Player',
   all.x = T
)

dtChartData[Player != 'Ball', Player := Label]
dtChartData[, Label := NULL]
   
   
   
p1 = ggplot(
  dtChartData
) + geom_pitch(cLineColour = '#aaaaaa', cPitchColour = NA) +
   geom_pitch(cLineColour = '#333333',cPitchColour = 'black') + 
   geom_path(
     data = dtChartData[
        Player != 'Ball'
     ][
     ][,
         # list(X = mean(X), Y = mean(Y)), list(Cluster, Player = gsub(Player, pattern = '.*Player', replacement = ''))
     ],
     aes(
       x = X,
       y = Y,
       group = paste(Instance, Cluster, Player)
     ),
     size = 0.2,
     color = 'red'
   ) +
   geom_point(
     data = dtChartData[
        Player != 'Ball'
     ][
     ][, .SD[which.max(Frame)], list(Player, Cluster, Instance)],
     aes(
       x = X,
       y = Y
     ),
     color = 'red',
     size = 0.2
   ) +
   geom_path(
     data = dtChartData[
        Player == 'Ball'
     ][
     ][,
         # list(X = mean(X), Y = mean(Y)), list(Cluster, Player = gsub(Player, pattern = '.*Player', replacement = ''))
     ],
     aes(
       x = X,
       y = Y,
       group = paste(Instance, Cluster, Player)
     ),
     size = 0.2,
     color = 'white'
   ) +
   geom_point(
     data = dtChartData[
        Player == 'Ball'
     ][
     ][, .SD[which.max(Frame)], list(Player, Cluster, Instance)],
     aes(
       x = X,
       y = Y
     ),
     size = 0.2,
     color = 'white'
   ) +
   # geom_polygon(
   #    data = dtChartDataHull,
   #    aes(
   #       x = X,
   #       y = Y,
   #       group = Cluster
   #    ),
   #    fill = NA,
   #    color = 'black'
   # ) +
   facet_wrap(~ClusterGroup.Cluster, nrow = 8) +
   scale_colour_discrete(guide = F) +
   geom_blank() + 
   coord_fixed() +
   theme_pitch() + 
   theme_pitch_override() +
   labs(
      title = paste0('Detailed positions of ', dtPlayerLabels[Player == cSelectedPlayer, Label], ' and ball across clusters'),
      caption = 'by @thecomeonman',
      subtitle = 'Ball positions in black, player positions in red.'
   )

print(p1)

```


## Cluster occurrence across the duration of the game
 
```{r PrintingClusterGroupsAcrossTime, cache=FALSE, echo=FALSE, warning=FALSE, message=FALSE, fig.width = 15, fig.height = 10}

dtSignificantEvents = rbind(
    dtSubstitutions[, list(Frame, Tag, Event = 'Substitution')],
    lData$dtEventsData[grepl(Subtype, pattern = 'GOAL$'), list(Frame = StartFrame, Tag = Team, Event = paste0(' Goal'))],
    lData$dtTrackingData[, max(Frame), list(Tag, Player)][V1 != max(V1), list(Tag, Frame = V1, Event = 'Red card')],
    lData$dtTrackingData[, list(Frame = max(Frame), Event = 'End of half', Tag = ''), Period][Period < max(Period), list(Tag, Frame, Event)]
)[, 
   list(Frame, Tag, Event, ClusterGroup = 'Events')
]

# vcAxisLabels = dtSignificantEvents[, paste0(Tag, ifelse(Tag == '','','\n'), Event)]
# names(vcAxisLabels) = dtSignificantEvents[, Frame * 0.04 / 60]

p1 = ggplot(dtClusterResults) +  
    geom_vline(
      data = dtSignificantEvents[, list(Frame)],
      aes(xintercept = Frame * 0.04 / 60),
      color = 'white',
      alpha = 0.5
    ) 

if ( F ) {
   
   p1 = p1 +  
      geom_text(
         data = dtSignificantEvents,
         aes(
            x = Frame * 0.04 / 60,
            y = 'Event',
            label = paste0(Tag, ifelse(Tag == '','','\n'), Event)
         ),
         color = 'white',
         alpha = 0.5
      )
   
}

if ( F ) {
   
   p1 = p1 +
       geom_point(
           data = dtSignificantEvents,
         aes(x = Frame * 0.04 / 60, y = paste(Tag, Event)),
         color = 'white'
       ) +  
       geom_hline(
           data = dtSignificantEvents,
         aes(yintercept = paste(Tag, Event)),
         alpha = 0.5,
         color = 'white'
       )
}

p1 = p1 + 
   geom_point(
      aes(x = Frame * 0.04 / 60, y = factor(Cluster), color = factor(ClusterGroup))
   )  + 
   geom_line(
      aes(x = Frame * 0.04 / 60, y = factor(Cluster), color = factor(ClusterGroup)), size = 0.1
   ) +
   facet_grid(
      ClusterGroup~., 
      scales = 'free', 
      space = 'free',
      labeller = label_both
   )

p1 = p1 +
   labs(
      title = 'Cluster occurrence across the duration of the game',
      x = 'Time',
      y = 'Cluster'
   ) +
   scale_color_discrete(guide = FALSE) +
   theme_non_pitch_override() +
   scale_x_continuous(
      breaks = c(
         dtSignificantEvents[, Frame * 0.04 / 60],
         c(15, 30),
         c(15, 30) + dtSignificantEvents[Event == 'End of half'][, Frame * 0.04 / 60]
      ),
      labels = c(
         dtSignificantEvents[, paste0('\n', Tag, ifelse(Tag == '','','\n'), Event)],
         paste0(c(15, 30), '\''),
         paste0(c(15, 30), '\'')
      )
   )

print(p1)

```
 
There are interesting patterns in terms of how clusters occur at various points in the game. After pretty much each significant event, you can notice a change in the patterns of clusters exhibited.


## The flow between clusters
 
 
```{r PrintingClusterTransitions1, cache=FALSE, echo=FALSE, warning=FALSE, message=FALSE, fig.width = 15, fig.height = 7}
   
dtTransitions = dtClusterResults[, 
   list(
      ClusterGroup = head(ClusterGroup, -1), 
      Cluster = head(Cluster, -1),
      Frame = head(Frame, -1), 
      NextClusterGroup = tail(ClusterGroup, -1), 
      NextCluster = tail(Cluster, -1),
      NextFrame = tail(Frame, -1)
   )
]

dtTransitions = dtTransitions[
   Cluster != NextCluster
][
   NextFrame - Frame < 10
][, 
   .N, 
   list(ClusterGroup, Cluster, NextClusterGroup, NextCluster)
]

   
if ( F ) {

   dtTransitionsGroup = dtTransitions[
      ClusterGroup != NextClusterGroup,
      list( N = sum(N) ), 
      list(ClusterGroup, NextClusterGroup)
   ]
   
   dtTransitionsGroup = merge(
      dtTransitionsGroup,
      dtTransitionsGroup[, list(NextClusterGroup = dtTransitionsGroup[, unique(ClusterGroup)]), ClusterGroup],
      c('ClusterGroup','NextClusterGroup'),
      all = T
   )
   
   dtTransitionsGroup = dtTransitionsGroup[ClusterGroup != NextClusterGroup]
   dtTransitionsGroup[is.na(N), N := 0]
   dtTransitionsGroup[, N_pct := N / sum(N), list(ClusterGroup)]
   dtTransitionsGroup[is.na(N_pct), N_pct := 0]
   
   dtTransitionsGroup = merge(
      dtTransitionsGroup,
      merge(
         dtClusterResults[!is.na(Cluster)],
         lData$dtTrackingData[Tag == 'Ball', list(Frame, X, Y)],
         'Frame'
      )[,
         list(
            X1 = mean(X),
            Y1 = mean(Y)
         ),
         list(
            ClusterGroup = ClusterGroup
         )
      ],
      'ClusterGroup'
   )
   
   dtTransitionsGroup = merge(
      dtTransitionsGroup,
      merge(
         dtClusterResults[!is.na(Cluster)],
         lData$dtTrackingData[Tag == 'Ball', list(Frame, X, Y)],
         'Frame'
      )[,
         list(
            X2 = mean(X),
            Y2 = mean(Y)
         ),
         list(
            NextClusterGroup = ClusterGroup
         )
      ],
      'NextClusterGroup'
   )
   
   if ( F ) {
         
      p1 = ggplot(dtTransitionsGroup) + 
         geom_tile(aes(x = factor(ClusterGroup), y = factor(NextClusterGroup), alpha = N_pct), fill = 'red') +
         coord_fixed() +
         scale_alpha_continuous(range = c(0, 1), limits = c(0, NA)) +
         labs(
            title = 'What cluster group occurs next',
            y = 'Next cluster group',
            x = 'Current cluster group'
         )
   
   } else {
      
      p1 = ggplot(dtTransitionsGroup) + 
         geom_pitch(cLineColour = '#333333', cPitchColour = 'black') + 
         geom_point(
            aes(
               x = X2,
               y = Y2,
               alpha = N_pct
            ),
            size = 20,
            color = 'white',
            fill = 'red',
            shape = 21
         ) + 
         geom_point(
            aes(
               x = X1,
               y = Y1
            ),
            color = 'white'
         ) + 
         geom_segment(
            aes(
               x = X1,
               y = Y1,
               xend = X2,
               yend = Y2
            ),
            color = 'white'
         ) + 
         scale_color_continuous(limits = c(0, 1), low = 'white', high = 'red', name = NULL) +
         scale_alpha_continuous(range = c(0, 1), labels = percent, name = NULL) +
         facet_wrap(~ClusterGroup) + 
         theme_pitch() + 
         theme_pitch_override() +
         labs(
            title = 'Movement between cluster groups'
         )
         
         
      if ( F ) {
         
         ggplot(dtTransitionsGroup) + 
            geom_pitch() + 
            geom_segment(
               aes(
                  x = X1,
                  y = Y1,
                  xend = X2,
                  yend = Y2,
                  # color = N_pct
               )
            ) + 
            geom_point(
               aes(
                  x = X2,
                  y = Y2,
                  size = N_pct
               )
            ) + 
            scale_color_continuous(limits = c(0, 1), low = 'black', high = 'red') +
            scale_size_continuous(range = c(0, 10), limits = c(0, 1)) +
            facet_wrap(~ClusterGroup)
         
      }
   
   }
   
   print(p1)

}

```


```{r PrintingClusterTransitions2, cache=FALSE, echo=FALSE, warning=FALSE, message=FALSE, fig.width = 15, fig.height = 7}




dtTransitions = merge(
   dtTransitions,
   data.table(expand.grid(
       Cluster = dtClusterResults[, unique(Cluster)],
       NextCluster = dtClusterResults[, unique(Cluster)]
   )),
   c('Cluster','NextCluster'),
   all = T
)

dtTransitions[, c('ClusterGroup','NextClusterGroup','ClusterGroup.Cluster','NextClusterGroup.Cluster') := NULL]

dtTransitions = merge(
   dtTransitions,
   dtClusterResults[, .SD[1], list(Cluster, ClusterGroup)][, list(Cluster, ClusterGroup, ClusterGroup.Cluster)],
   'Cluster'
)
dtTransitions = merge(
   dtTransitions,
   dtClusterResults[, .SD[1], list(Cluster, ClusterGroup)][, list(NextCluster = Cluster, NextClusterGroup = ClusterGroup, NextClusterGroup.Cluster = ClusterGroup.Cluster)],
   'NextCluster'
)

dtTransitions = dtTransitions[Cluster != NextCluster]
dtTransitions[is.na(N), N := 0]
dtTransitions[, N_pct := N / sum(N), list(Cluster)]

dtTransitions = merge(
   dtTransitions,
   merge(
      dtClusterResults[!is.na(Cluster)],
      lData$dtTrackingData[Tag == 'Ball', list(Frame, X, Y)],
      'Frame'
   )[,
      list(
         X1 = mean(X),
         Y1 = mean(Y)
      ),
      list(
         Cluster = Cluster
      )
   ],
   'Cluster'
)

dtTransitions = merge(
   dtTransitions,
   merge(
      dtClusterResults[!is.na(Cluster)],
      lData$dtTrackingData[Tag == 'Ball', list(Frame, X, Y)],
      'Frame'
   )[,
      list(
         X2 = mean(X),
         Y2 = mean(Y)
      ),
      list(
         NextCluster = Cluster
      )
   ],
   'NextCluster'
)

dtTransitions[is.na(N_pct), N_pct := 0]

p1 = ggplot(dtTransitions) + 
   geom_pitch(cLineColour = '#333333',cPitchColour = 'black') + 
   geom_segment(
       aes(x = X1, y = Y1, xend = X2, yend = Y2, alpha = N), 
       arrow = arrow(), 
       color = 'red'
   ) +
   geom_point(
      data = merge(
         dtClusterResults[, 
            list(
               Frame, 
               Cluster, 
               ClusterGroup, 
               Instance
            )
         ], 
         lData$dtTrackingData[
            Tag == 'Ball', 
            list(X, Y, Frame)
         ],
         c('Frame')
      )[, 
         list(
            X = mean(X), 
            Y = mean(Y), 
            Instance = unique(Instance)
         ), 
         list(
            ClusterGroup,
            Cluster
         )
      ], 
      aes(
         x = X, 
         y = Y, 
         size = Instance
      ), 
      fill = 'red',
      color = 'black',
      shape = 21
   ) +
    geom_point(
       data = dtTransitions[, list(N = sum(N)), list(X2, Y2, NextCluster, ClusterGroup)][N > 0], 
      aes(
         x = X2, 
         y = Y2, 
         size = N
      ), 
      fill = 'white',
      color = 'black',
      shape = 21
   ) +
   scale_alpha_continuous(range = c(0, 1), name = 'Transitions\nbetween\nclusters') +
   scale_size_continuous(range = c(0, 5), limits = c(0, NA), name = 'Instances\nof this\ncluster') +
   theme_pitch() + 
   theme_pitch_override() +
   labs(
      title = 'Movement between clusters'
   ) + 
   facet_wrap(~ClusterGroup)

print(p1)

if ( F ) {
   
   p1 = ggplot(dtTransitions) + 
      geom_pitch(cLineColour = '#333333',cPitchColour = 'black') + 
      geom_point(
         aes(
            x = X2,
            y = Y2,
            alpha = N_pct
         ),
         size = 5,
         color = 'white',
         fill = 'red',
         shape = 21
      ) + 
      geom_point(
         aes(
            x = X1,
            y = Y1
         ),
         color = 'white'
      ) + 
      # geom_segment(
      #    aes(
      #       x = X1,
      #       y = Y1,
      #       xend = X2,
      #       yend = Y2
      #    ),
      #    color = 'white'
      # ) + 
      scale_color_continuous(limits = c(0, 1), low = 'white', high = 'red', name = NULL) +
      scale_alpha_continuous(range = c(0, 1), labels = percent, name = NULL) +
      facet_wrap(~ClusterGroup.Cluster, nrow = 6) + 
      theme_pitch() + 
      theme_pitch_override()
     
   print(p1)

}
```
 
The team doesn't randomly move between clusters, there are some transitions that occur far more frequently than others.
 
