---
title: "Novel Uses of Tracking Data in Modern Football"
date: "30/05/2020"
output: md_document
---

## Before you begin:

Before attempting to run this code, follow the instructions at 
https://github.com/thecomeonman/CodaBonito to install some dependencies.

Also get the tracking data, recently made public as part of the Friends of 
Tracking sessions by Metrica from here - https://github.com/metrica-sports/sample-data

## The code

```{r Setup, cache=FALSE, echo=FALSE, warning=FALSE, message=FALSE, fig.width = 15, fig.height = 8, results = 'hide'}

library(CodaBonito)
library(data.table)
library(ggplot2)
library(ggrepel)
theme_set(theme_bw(12))

rm(list = ls())

# Instructions
# You would have downloaded and saved the data from Github at some location on
# your system. In my case, it's /media/ask/Data/Personal/Projects/Personal/sample-data
# Replace the location below with the respective location from your machine
cDataRootFolder = '/media/ask/Data/Personal/Projects/Personal/sample-data/data/'
cGameName = 'Sample_Game_1'
nXLimit = 120
nYLimit = 80
# upper limit of speed to cap weird speed values
nUpperLimitSpeed = 5 * 120 / 105

```

```{r DataLoading, cache=FALSE, echo=FALSE, warning=FALSE, message=FALSE, fig.width = 15, fig.height = 8, results = 'hide'}

# loading the data in
lData = fParseTrackingDataBothTeams(
   cRootPath = cDataRootFolder,
   cGameName = cGameName,
   nXLimit = nXLimit,
   nYLimit = nYLimit,
   xMaxBB = 1,
   yMaxBB = 1,
   nUpperLimitSpeed = nUpperLimitSpeed
)

```


```{r CalculatingDistances, cache=FALSE, echo=FALSE, warning=FALSE, message=FALSE, fig.width = 15, fig.height = 8, results = 'hide'}

dtSocialDistancing = rbindlist(
   lapply(
      lData$dtTrackingData[, seq(ceiling(max(Frame) / 100))],
      function( iFrame ) {
         
         print(iFrame)
         
         dtSubset = lData$dtTrackingData[
            Frame %in% (1 + (100 * (iFrame-1))):(100 * (iFrame))
         ][
            Player != 'Ball'
         ]
         
         dtSubset = merge(
            dtSubset[, list(Frame, Player, X, Y)],
            dtSubset[, list(Frame, Player, X, Y)],
            'Frame',
            allow.cartesian = T
         )
         
         dtSubset = dtSubset[Player.x != Player.y]
         
         dtSubset[, 
            Distance := (
               ( ( X.x - X.y ) ^ 2 ) +
               ( ( Y.x - Y.y ) ^ 2 )
            ) ^ 0.5
         ]
         
         dtSubset = dtSubset[Distance < 7]
         
         dtSubset
         
      }
      
   )
)


setorder(
   dtSocialDistancing,
   Player.x,
   Player.y,
   Frame
)

dtSocialDistancing[, Stretch2 := (c(1, diff(Frame) > 25)), list(Player.x, Player.y)]
dtSocialDistancing[, Stretch := cumsum(c(1, diff(Frame) > 25)), list(Player.x, Player.y)]

```

```{r TotalTimeSpent, cache=FALSE, echo=FALSE, warning=FALSE, message=FALSE, fig.width = 15, fig.height = 8, results = 'hide'}

ggplot(
   dtSocialDistancing[, list(TimeSpentWithin6Feet = .N * 0.04), list(Player.x, Player.y)]
) + 
   geom_tile(
      aes(
         x = Player.x,
         y = Player.y,
         alpha = TimeSpentWithin6Feet / 60
      ),
      fill = 'red',
      color = 'black'
   ) +
   geom_text(
      aes(
         x = Player.x,
         y = Player.y,
         label = round(TimeSpentWithin6Feet / 60)
      )
   ) +
   coord_fixed() +
   theme(
      axis.text.x = element_text(angle = 90)
   ) +
   scale_alpha_continuous(guide = F) +
   labs(
      title = 'Minutes spent within ~6 feet of each other ',
      x = NULL,
      y = NULL,
      caption = '- @thecomeonman'
   )

```

```{r TotalOccurrences, cache=FALSE, echo=FALSE, warning=FALSE, message=FALSE, fig.width = 15, fig.height = 8, results = 'hide'}

ggplot(
   dtSocialDistancing[, list(Occurrence = length(unique(Stretch))), list(Player.x, Player.y)]
) + 
   geom_tile(
      aes(
         x = Player.x,
         y = Player.y,
         alpha = Occurrence
      ),
      fill = 'red',
      color = 'black'
   ) +
   geom_text(
      aes(
         x = Player.x,
         y = Player.y,
         label = Occurrence
      ),
      size = 2
   ) +
   coord_fixed() +
   theme(
      axis.text.x = element_text(angle = 90)
   ) +
   scale_alpha_continuous(guide = F) +
   labs(
      title = 'Instances of being within ~6 feet of each other ',
      x = NULL,
      y = NULL,
      caption = '- @thecomeonman'
   )

```

```{r TotalTimeSpentContinuous, cache=FALSE, echo=FALSE, warning=FALSE, message=FALSE, fig.width = 15, fig.height = 8, results = 'hide'}

ggplot(
   dtSocialDistancing[, list(Duration = .N * 0.04), list(Stretch, Player.x, Player.y)]
) + 
   geom_jitter(aes(x = Player.x, y = Duration / 60), width = 0.2) +
   geom_text_repel(
      data = dtSocialDistancing[, list(Duration = .N * 0.04), list(Stretch, Player.x, Player.y)][Duration > 60],
      aes(
         x = Player.x,
         y = Duration / 60,
         label = Player.y
      )
   ) +
   labs(
      title = 'Minutes spent within ~6 feet of any other player',
      x = NULL,
      y = NULL,
      caption = '- @thecomeonman'
   ) +
   theme(
      axis.text.x = element_text(angle = 90)
   )


```

```{r LongestStretch, cache=FALSE, echo=FALSE, warning=FALSE, message=FALSE, fig.width = 15, fig.height = 8, results = 'hide'}


dtLongestStretch = dtSocialDistancing[, list(Duration = .N * 0.04), list(Stretch, Player.x, Player.y)][which.max(Duration), list(Player.x, Player.y, Stretch)]

p1 = ggplot(
   lData$dtTrackingData[
      Frame %in% merge(
         dtLongestStretch,
         dtSocialDistancing,
         c('Player.x', 'Player.y', 'Stretch')
      )[, Frame]
   ][
      Player %in% dtLongestStretch[, c(Player.x, Player.y)]
   ]
) + 
   geom_path(aes(x = X, y = Y, group = Player, color = Player)) +
   labs(
      title = merge(
         dtLongestStretch,
         dtSocialDistancing,
         c('Player.x', 'Player.y', 'Stretch')
      )[, paste0('Longest stretch of staying in each other\'s vicinty: from frame ', min(Frame), ' to frame ', max(Frame))],
      caption = '- @thecomeonman'
   )

p1 = fAddPitchLines(
   p1,
   nXLimit = nXLimit, 
   nYLimit = nYLimit, 
   cLineColour = 'black', 
   cPitchColour = NA
) +
   theme_pitch()

p1
